---
title: "Differences in U.S. Expenditure Pay for Recreational Workers in States of America from 1930-1959 and its Effect on Population and Culture"
name: David P. Askew
date: 5/1/2024
format: html
editor: visual
---

### David P. Askew

#For the final project, I am using digital tools like data visualization and topic modeling to examine, compare, and contrast expenditure pay in different regions of the United States during the 1930s up to 1959.

## What is the Purpose of This Project?

This project will answer a few research questions I have:

1.  What is the difference in pay between recreational workers in 1930s-1950s New England and Southern states and why did this difference exist based on prominence of recreational potential?

2.  How does the difference in pay affect the population of people in these states? What can it tell us about the effect it has on cultures between regions in the United States?

## Introduction

During the 1930s, the North and South were starkly different regions of the United States as they existed at the time with different populations. This certainly applies to the difference in pay based on the amount of recreational activities between states and how much of it applied to an economic contrast. The major differences between the North and South culturally can still be felt to this day. However it is important recognize the way which the 1930s to 1950s, through recreation, effectively shaped both regions economically and with regard to population.

Through using the tools of data visualization and topic modeling, this analysis will aim to explore how recreation during the 1930s-1950s had an impact on both pay and the population of Northern and Southern states. By examining the prominence of recreation in different states in the 1930s-1950s based on population data, we can better understand the differences in pay for recreational workers in those states. The second half of this research will observe how the difference in pay affected the populations of those states in different regions and the effect it had on their cultures.

## Historical Background

Recreation had a tremendous impact on workers in many U.S. states and effectively shaped their culture. This project analyzes and compares the differences between Northern and Southern states related to pay toward recreational workers in the 1930s-1950s based on prominence of recreation and impact it has on population of people in those states as well as their cultures. According to Elaine Tyler May's *Homeward Bound: American Families in the Cold War Era,* May explains that during the 1930s, recreation had gradually increased for many Americans leading up to the 1940s. This was primarily because the economic crisis brought on by the Great Depression left many in need for jobs which is where FDR's New Deal would attempt to resolve this issue.[^1] The New Deal was effective in ways that despite eliminating poverty or solving the economic crisis, helped to ease suffering by giving federal occupations to people in need of work. These would generally be through youth programs like the Civilian Conservation Corps where young, single Americans would work at camps for \$30 a month.[^2] May also states that the income was $2,000 a year for 

[^1]: Elaine Tyler May, *Homeward Bound: American Families in the Cold War Era* (New York: Basic Books, 2008), 41.

[^2]: National Park Service. \"Civilian Conservation Corps - Theodore Roosevelt National Park (U.S. National Park Service).\" Nps.gov. 2016. https://www.nps.gov/thro/learn/historyculture/civilian-conservation-corps.htm. Accessed May 1, 2024.

    ‌

```{r}
library(tm)
library(topicmodels)
library(tidytext)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tidyverse)
library(word2vec)

```

```{r}
sc.rec <- recreational.data %>% group_by(state) %>% summarize(state.expenditures = sum(total_expenditures))
ggplot(sc.rec, aes(x = state, y = state.expenditures )) +
  geom_col() + ggtitle("Differences in Expenditure Between States")
```

```{r}
rec.byyear <- recreational.data %>% filter(state == "CT" | state == "NY" | state == "SC") %>% group_by(state, year) %>% summarize(count = n())
ggplot(rec.byyear, aes(x = state, y = count)) +
  geom_bar()
```

```{r}

```

```{r}
rec.workers <- read.csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/RecreationData-Wide.csv")
rec.workers.narrow <- rec.workers %>% pivot_longer(!city:type_of_worker, names_to = "year", values_to = "count")
  all.paid <- c("male paid", "female paid")
    all.volunteer <- c("men volunteer", "women volunteer")
      avg.volunteer <- rec.workers.narrow %>% group_by(year) %>% filter(type_of_worker %in% all.volunteer) %>% summarise(avg.volunteer.count = mean(count, na.rm = TRUE))
        avg.paid <- rec.workers.narrow %>%  group_by(year) %>% filter(type_of_worker %in% all.paid) %>% summarise(avg.paid.count = mean(count, na.rm = TRUE))
            paid.and.volunteer <- merge(avg.volunteer, avg.paid, by = "year")
              rec.workers.longer <- pivot_longer(paid.and.volunteer, !year, names_to = "type.of.workers", values_to = "count")
ggplot(data = rec.workers.longer, mapping = aes(x = year, y= count, group = type.of.workers, color = type.of.workers)) + geom_line() + labs( x = "Year", y= "Number of Workers", Title = "Volunteer and Paid Rec Workers Over Time", color = "Type of Workers") + theme(plot.title = element_text(hjust = 0.5)) + theme_dark() + geom_point()
```

## How Does the Difference in Pay for Recreational Workers in the North and South Relate to Cultural Division in Post World War II America?
