---
title: "Differences in U.S. Expenditure for Recreational Workers in the United States from 1930-1947 and its Effect on Population, Culture, and Geography"
name: David P. Askew
date: 5/1/2024
format: html
editor: visual
---

### David P. Askew

For the final project, I am using the digital tool of data visualization to examine, compare, and contrast recreational expenditure pay in different regions of the United States with that of population of states during the 1930s.

## What is the Purpose of This Project?

This project will answer a few research questions I have:

1.  What is the difference in expenditures between recreational workers in Northern and Southern states during the 1930s and why did this difference exist based on prominence of recreational potential?

2.  How does the difference in expenditure affect the population of people in these states? What can it tell us about the effect it has on cultures between regions in the United States?

# Introduction

During the 1930s, the North and South were starkly different regions of the United States as they existed at the time with different populations. This certainly applies to the difference in pay based on the amount of recreational activities between states and how much of it applied to an economic contrast. The major differences between the North and South culturally can still be felt to this day. However it is important recognize the way which the 1930s to the late 1940s, through recreation, effectively shaped both regions economically and with regard to population. Furthermore, it is important to see how racial and sexual discrimination played a role in encouraging the environment of these states.

Through using the tools of data visualization, this analysis will aim to explore how recreation during the 1930s had an impact on both pay and the population of Northern and Southern states. By examining the prominence of recreation in different states in the 1930s based on population data, we can better understand the differences in pay for recreational workers in those states. The second half of this research will observe how the difference in pay affected the populations of those states in different regions during the 1930s and the effect it had on their economy and geography.

## Historical Background

Recreation had a tremendous impact on workers in many U.S. states and effectively shaped their culture. This project analyzes and compares the differences between Northern and Southern states related to pay toward recreational workers in the 1930s based on prominence of recreation and impact it has on population of people in those states as well as their cultures. According to Elaine Tyler May's *Homeward Bound: American Families in the Cold War Era,* May explains that during the 1930s, recreation had gradually increased for many Americans leading up to the 1940s. This was primarily because the economic crisis brought on by the Great Depression left many in need for jobs which is where FDR's New Deal would attempt to resolve this issue.[^1] The New Deal was effective in ways that despite eliminating poverty or solving the economic crisis, helped to ease suffering by giving federal occupations to people in need of work.

[^1]: Elaine Tyler May, *Homeward Bound: American Families in the Cold War Era* (New York: Basic Books, 2008), 41.

These would generally be through youth volunteer programs like the Civilian Conservation Corps where young, single Americans would work at camps for \$30 a month.[^2] May also states that the median income for most families was \$2,000 a year by 1941 and many movies during the 1930s emphasized dealing with the hardship of the Depression through the equal roles of men and women at work.[^3] There was a collective effort for families to overcome the hardship through labor and this was apparent in many cultural aspects of the United States at the time. By looking at the difference of expenditures between states, we can understand how impactful this period was.

[^2]: National Park Service. "Civilian Conservation Corps - Theodore Roosevelt National Park (U.S. National Park Service)." Nps.gov. 2016. https://www.nps.gov/thro/learn/historyculture/civilian-conservation-corps.htm. Accessed May 1, 2024.

[^3]: May, 43.

In Thomas Maloney's "Wage Compression and Wage Inequality Between Black and White Males in the United States", it can be recognized how the South has a lower expenditure because of discrimination from these workers being mostly African American.[^4] Maloney explains wages were lower in the South than anywhere else during that period stating, "Black migration to cities both in the South and in other regions caused the proportion urban among blacks to exceed that among whites for the first time by 1960."[^5] This is a sharp contrast between that of states in the North as explained by Lawrence Chidester which had a well-established vacation industry for visitors like hotels, lodging, restaurants, and campgrounds which allowed for more job opportunities and higher wages.[^6]

[^4]: Thomas N. Maloney, "Wage Compression and Wage Inequality Between Black and White Males in the United States, 1940-1960." *The Journal of Economic History* 54, no. 2 (1994): 358--81. http://www.jstor.org/stable/2123918. Accessed May 2, 2024.

[^5]: Maloney, 367.

[^6]: Lawrence W. Chidester, "The Importance of Recreation as a Land Use in New England." *The Journal of Land & Public Utility Economics* 10, no. 2 (1934): 202--9. https://doi.org/10.2307/3138921. Accessed May 2, 2024.

## Data Visualization ‌

```{r}
#1 Loading libraries

library(tidytext)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tidyverse)
```

```{r}
#2 Reading data

recreational.data <- readr::read_csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/Recreation-Expenditures.csv")
```

```{r}
#3 Creating a bar plot that demonstrates total differences in expenditure between states during the 1930s

sc.rec <- recreational.data %>% group_by(state) %>% summarize(state.expenditures = sum(total_expenditures))
ggplot(sc.rec, aes(x = state, y = state.expenditures )) +
  geom_col() + ggtitle("Total Differences in Expenditure Between States in the 1930s")
```

```{r}
#3 Creating a bar plot for total expenditure in South Carolina during the 1930s

library(dplyr)
library(ggplot2)
sc.rec <- recreational.data %>%
  filter(state == "SC") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(sc.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for South Carolina in the 1930s")
```

```{r}
#3 Creating a bar plot for total expenditure in Tennessee during the 1930s

library(dplyr)
library(ggplot2)
tn.rec <- recreational.data %>%
  filter(state == "TN") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(tn.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for Tennessee in the 1930s")
```

```{r}
#4 Creating a bar plot for total expenditure in Texas during the 1930s

library(dplyr)
library(ggplot2)
tx.rec <- recreational.data %>%
  filter(state == "TX") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(tx.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for Texas in the 1930s")
```

```{r}
#4 Creating a line graph for total expenditures in the South during the 1930s (SC, TN, TX)

library(dplyr)
library(ggplot2)
rec.data <- recreational.data %>%
  filter(state %in% c("SC", "TN", "TX")) %>%
  group_by(state, year) %>%
  summarize(state.expenditures = sum(total_expenditures))
print(ggplot(rec.data, aes(x = year, y = state.expenditures, color = state, group = state)) + geom_line() + ggtitle("Total Expenditures for the South Over Time in the 1930s"))
```

```{r}
#5 Creating a bar plot for total expenditure in South Dakota during the 1930s

library(dplyr)
library(ggplot2)
sd.rec <- recreational.data %>%
  filter(state == "SD") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(sd.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for South Dakota in the 1930s")
```

```{r}
#6 Creating a bar plot for total expenditure in Rhode Island during the 1930s

library(dplyr)
library(ggplot2)
ri.rec <- recreational.data %>%
  filter(state == "RI") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(ri.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for Rhode Island in the 1930s")
```

```{r}
#6 Creating a bar plot for total expenditure in Pennsylvania during the 1930s

library(dplyr)
library(ggplot2)
pa.rec <- recreational.data %>%
  filter(state == "PA") %>%
  group_by(state) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(pa.rec, aes(x = state, y = state.expenditures)) +
  geom_col() +
  ggtitle("Total Expenditure for Pennsylvania in the 1930s")
```

```{r}
#7 Creating a line graph for total expenditures in the North during the 1930s (SD, RI, and PA)

library(dplyr)
library(ggplot2)
rec.data <- recreational.data %>%
  filter(state %in% c("SD", "RI", "PA")) %>%
  group_by(state, year) %>%
  summarize(state.expenditures = sum(total_expenditures))
ggplot(rec.data, aes(x = year, y = state.expenditures, color = state, group = state)) +
  geom_line() +
  ggtitle("Total Expenditures for the North Over Time in the 1930s")
```

```{r}
#8 Creating a line graph that shows the average paid recreational workers compared with that of volunteers over time from the 1930s to 1945.

rec.workers <- read.csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/RecreationData-Wide.csv")
rec.workers.narrow <- rec.workers %>% pivot_longer(!city:type_of_worker, names_to = "year", values_to = "count")
  all.paid <- c("male paid", "female paid")
    all.volunteer <- c("men volunteer", "women volunteer")
      avg.volunteer <- rec.workers.narrow %>% group_by(year) %>% filter(type_of_worker %in% all.volunteer) %>% summarise(avg.volunteer.count = mean(count, na.rm = TRUE))
        avg.paid <- rec.workers.narrow %>%  group_by(year) %>% filter(type_of_worker %in% all.paid) %>% summarise(avg.paid.count = mean(count, na.rm = TRUE))
            paid.and.volunteer <- merge(avg.volunteer, avg.paid, by = "year")
              rec.workers.longer <- pivot_longer(paid.and.volunteer, !year, names_to = "type.of.workers", values_to = "count")
ggplot(data = rec.workers.longer, mapping = aes(x = year, y= count, group = type.of.workers, color = type.of.workers)) + geom_line() + labs( x = "Year", y= "Number of Workers", Title = "Volunteer and Paid Rec Workers Over Time", color = "Type of Workers") + theme(plot.title = element_text(hjust = 0.5)) + theme_dark() + geom_point()
```

## Findings

My analysis on the difference in expenditures for Northern and Southern states during the 1930s as well as their impact on recreational workers demonstrates as time went on, the number of workers gradually increased, and the economy improved into the 1940s. Additionally, the data suggests that Southern states, with the exception of Texas, paid lower expenditures than the North. This had to deal with differences in geography particularly regarding urban vs. rural landscapes and whether these states were affected as badly by the Dust Bowl.

The reason why Texas stood out in terms of expenditure was because during the Great Depression, according to the Texas State Historical Association, the state was the epicenter of the Civilian Conservation Corps programs.[^7] In the article "Great Depression" by Ben Procter and Brian Cervantez, "The CCC, a nationwide program, for young men that focused on natural resources from 1933 to 1942, was very active in Texas. At its peak in 1935 the corps had twenty seven camps in Texas constructing recreational parks and an additional seventy camps for work in forest and soil conservation."[^8] This also deals with the fact that Texas was one of the states in the Dust Bowl and this affected it greatly. The desperation caused by these hard times manifested in the form of recreation. Therefore, based on this secondary source and the data demonstrated by my analysis, Texas was the state with the most potential for recreation in the South which led to higher expenditures.

[^7]: Ben H. Procter; Brian A. Cervantez, "Great Depression," *Handbook of Texas Online,* October 1, 1995, Accessed May 02, 2024, https://www.tshaonline.org/handbook/entries/great-depression.

[^8]: Procter; Cervantez, "Great Depression".

When looking at the North, the highest expenditure was in Pennsylvania for similar reasons as with Texas in the South. As explained by the *Pennsylvania Department of Conservation and Natural Resources*, Pennsylvania was the state with the most potential in the North for its parks, forests, and mountains.[^9] According to the website, "From 1933 to 1942, hundreds of thousands of unemployed men worked in camps throughout Pennsylvania (...) Some camps were on army bases, some were in national parks \[...\] but most camps were in state parks and forests. \[...\] Pennsylvania had the second highest number of camps (151) to California."[^10] The outlier in the data was South Dakota which had the lowest expenditure out of the rest of the other states.

[^9]: Pennsylvania Department of Conservation and Natural Resources, "The Civilian Conservation Corps", May 2, 2024. [CCC Years (pa.gov)](https://www.dcnr.pa.gov/StateParks/History/CCCYears/Pages/default.aspx). Accessed May 2, 2024.

[^10]: Pennsylvania Department of Conservation and Natural Resources, "CCC in Pennsylvania," [CCC Years (pa.gov)](https://www.dcnr.pa.gov/StateParks/History/CCCYears/Pages/default.aspx).

According to "How Reading Got Farm Women Through the Depression" by Livia Gershon, she states that "more than 80 percent of South Dakota's population lived in rural areas in the 1930s. The households suffered immensely from drought, grasshopper infestations, and falling crop prices. \[...\] reading seemed like a justifiable form of leisure."[^11] This is significant because despite initial assumptions the state would have much more potential because of its size, the state had little to no money during the Great Depression, which did not easily allow for recreation.

[^11]: Livia Gershon, "How Reading Got Farm Women Through the Depression," *JSTOR.* May 4, 2020. [How Reading Got Farm Women Through the Depression - JSTOR Daily](https://daily.jstor.org/how-reading-got-farm-women-through-the-depression/). Accessed May 5, 2024.

### This is the second part of my project where I will use data visualization to understand the effect that difference in expenditures had on populations of these states

```{r}
#1 Loading my libraries

library(tidytext)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tidyverse)
```

```{r}
#2 Finding the population of different states for the recreational.data

library(dplyr)
library(ggplot2)
populations <- recreational.data %>%
  group_by(state) %>%
  summarize(total_population = sum(population))
ggplot(populations, aes(x = state, y = total_population)) +
  geom_col() +
  ggtitle("Population in Each State in the 1930s")


```

```{r}
#3 Creating a line graph to demonstrate the relationship between population and expenditures of U.S. states

data_long_ordered <- data_long %>%
  arrange(state)
ggplot(data_long_ordered, aes(x = state, y = value, color = variable, group = variable)) +
  geom_path() +
  labs(y = "Value", title = "Population and Expenditures in Each State", color = "Variable")
```

```{r}
#4 Scatterplot showing population in the South in the 1930s

populations_filtered <- populations %>%
  filter(state %in% c("SC", "TN", "TX"))
ggplot(populations_filtered, aes(x = state, y = total_population)) +
  geom_point() +
  ggtitle("Population in the South in the 1930s")
```

```{r}
#5 Scatterplot showing population in the North in the 1930s

populations_filtered <- populations %>%
  filter(state %in% c("SD", "RI", "PA"))
ggplot(populations_filtered, aes(x = state, y = total_population)) +
  geom_point() +
  ggtitle("Population in the North in the 1930s")
```

### How Does the Difference in Expenditure for Recreational Workers in the North and South Relate to Population?

Given these results, expenditure of recreational workers relates very much to state population during the 1930s because of various political and economic circumstances brought on by the Great Depression. President Roosevelt's New Deal was a significant factor because it allowed for the expansion of federal jobs which led to an increase in populations for states where recreational activity was most apparent. This includes states like Pennsylvania and Texas. In fact, the article "Great Depression" in *The* *Encyclopedia of Greater Philadelphia* by Roger D. Simon states that the relief programs aided the population very much in a way which would encourage growth.[^12]

[^12]: Roger D. Simon, "Great Depression," *The Encyclopedia of Greater Philadelphia.* 2013. [Great Depression - Encyclopedia of Greater Philadelphia (philadelphiaencyclopedia.org)](https://philadelphiaencyclopedia.org/essays/great-depression/). Accessed May 2, 2024.

Population was very much affected in different states by however much the government funded recreational programs for each one. It also depended on the geography of a state which could allow for recreation to thrive. For example, states like South Dakota had lower levels of people which could not translate into enough recreational potential while states with more people like Texas had higher amounts of expenditure pay. Additionally, after doing my analysis, whether a state is rural or urban does not seem to matter as much as the population density. While it is important to take into consideration that many rural states did have less expenditure, their population was mildly larger with the exception of South Dakota. These include states such as Tennessee and South Carolina.

On the other hand, states that could be considered more urban like Rhode Island had less expenditure and a slightly higher population than both South Carolina and South Dakota. Why is this the case? The answer is because, like previously explained, this state was seen by many as a vacation destination. Even though there was not as much expenditure related to recreation, it had a higher population because of its appeal to tourists and people looking to go on vacation.

## Conclusion

It is intriguing to see from my analysis the way which populations fluctuated in different states, examining the rural vs. urban divide based on recreational expenditure, geography of different regions, and recreational activity in each state. Recognizing the difference in populations of Southern vs Northern states makes sense when considering the divide between rural and urban geography. The rural states like South Carolina and South Dakota have a lower population density because they received less government funding whereas states that received more, like Pennsylvania and Texas, had higher densities. It is important to note Pennsylvania is not considered an urban state, however this research has helped me understand that selective government funding brought on by the New Deal was what really impacted both the economy and population of these states. Moving forward as we reflect on the 1930s, it is important to see how the significance of rural vs urban geography was not as eventful in the bigger picture with population density and expenditures. Nevertheless, it did play a role in influencing those populations based on appeal.

## Bibliography

\[1\]
